Univariate model using Wilcox Rank Sum test

### 2020-01-10 scripts/31_univariate_lm_wilcox.Rmd
update

@@

```{r}
## Filter to have each pathway with at least 10% of nonzero (zero <= 90%)

#num_zero_thre <- floor(ncol(pw_fil) * 0.9)

#pw_fil <- pw_fil[rowSums(pw_fil == 0) <= num_zero_thre, ]

##### How many pathways left
#nrow(pw_fil) 
```

## Log2 transform the data

```{r}
#pw_Fil <- log2(pw_fil + 0.01)

# use the whole table (not filter anything)
pw_log2 <- log2(pw + 0.01)
```

How many pathways went into the test

```{r}
nrow(pw_log2)
```

## Wilcox for each pathway with Toxicity and CR as response respectively, followed by FDR correction

```{r}
phe_tox <- final_pheno$Toxicity
phe_CR <- final_pheno$CR

res_tox <- apply(pw_log2, 1, function(r){
  res = wilcox.test(r ~ phe_tox)
  return(pval = res$p.value)
}) %>% 
  data.frame() %>% 
  rename(pval = names(.)[1]) %>% 
  rownames_to_column('pw') %>% 
  mutate(padj = p.adjust(pval, method = 'BH')) %>% 
  arrange(desc(padj))

res_CR <- apply(pw_log2, 1, function(r){
  res = wilcox.test(r ~ phe_CR)
  return(pval = res$p.value)
}) %>% 
  data.frame() %>% 
  rename(pval = names(.)[1]) %>% 
  rownames_to_column('pw') %>% 
  mutate(padj = p.adjust(pval, method = 'BH')) %>% 
  arrange(desc(padj))

res_tox %>% 
  kable(caption = 'Response to Toxicity') %>% 
  kable_styling(full_width = F, position = 'left')

res_CR %>% 
  kable(caption = 'Response to CR') %>% 
  kable_styling(full_width = F, position = 'left')
```

## How many pathways are significant with FDR < 0.1

```{r}
# Toxicity
res_tox %>% 
  filter(padj < 0.1) %>% 
  nrow

# CR
res_CR %>% 
  filter(padj < 0.1) %>% 
  nrow

```

>>>>

phe_tox <- final_pheno$Toxicity
phe_CR <- final_pheno$CR

@@

## Use different prevalence filtering threshold (percentage of zero count samples) and do the log2 transform and wilcox test

```{r}
num_zero_thre <- seq(0, 0.99, 0.05)

test_diff_zero_thre_filter <- function(fil_thre){

  num_zero_thre <- floor(ncol(pw_fil) * fil_thre)

  pw_fil <- pw[rowSums(pw == 0) <= num_zero_thre, ]

  pw_log2 <- log2(pw_fil + 0.01)

  sig_tox <- apply(pw_log2, 1, function(r){
      res = wilcox.test(r ~ phe_tox, exact = F)
      return(pval = res$p.value)
    }) %>% 
      data.frame() %>% 
      rename(pval = names(.)[1]) %>% 
      rownames_to_column('pw') %>% 
      mutate(padj = p.adjust(pval, method = 'BH')) %>% 
      arrange(desc(padj)) %>% 
      filter(padj < 0.25) %>% 
      nrow

  sig_CR <- apply(pw_log2, 1, function(r){
      res = wilcox.test(r ~ phe_CR, exact = F)
      return(pval = res$p.value)
    }) %>% 
      data.frame() %>% 
      rename(pval = names(.)[1]) %>% 
      rownames_to_column('pw') %>% 
      mutate(padj = p.adjust(pval, method = 'BH')) %>% 
      arrange(desc(padj)) %>% 
      filter(padj < 0.25) %>% 
      nrow

  return(list(sig_tox_num = sig_tox,
         sig_CR_num = sig_CR,
         pw_num = nrow(pw_log2)))
}

res <- list()
for (i in 1:length(num_zero_thre)){
  res[[i]] = test_diff_zero_thre_filter(num_zero_thre[i])
}

final <- bind_rows(res) %>% 
  mutate(zero_perc = num_zero_thre) %>% 
  select(zero_perc, pw_num, sig_tox_num, sig_CR_num)

final %>% 
  kable(caption = 'Number of significant results (FDR < 0.25) in different zero percentage threshold') %>% 
  kable_styling(full_width = F, position = 'left')
```
